---
title: "R codes for Table 2 and Table 3"
output: html_notebook
---

R version 4.3.1 (2023-06-16) -- "Beagle Scouts"

Platform: aarch64-apple-darwin20 (64-bit)

ATTENTION: Neural networks leverage randomness to allow the architecture to effectively learn the function being approximated — they are stochastic by design. This means the same neural network can be trained on the same data but still produce different results each time the code is run.

#0.environment
```{r}
library(forecast)
library(tseries)
library(prophet)
set.seed(10)
Mape <- function(forecast, test)  {mean(abs( (forecast - test)/test ))}
Rmse <- function(forecast, test)  {sqrt(mean((forecast - test)^2))}
```

#1.Table 2
```{r}
# raw data
ts_data <- openxlsx::read.xlsx("./raw data for table 2:3.xlsx", sheet = 1, rows = 38:160, cols = 3)
ts_data <- ts(ts_data, frequency = 12, start = c(2013,1))
train_data <- window(ts_data, end = c(2018,12))
test_data <- window(ts_data, start = c(2019,1), end = c(2019,12))

# modeling
# Arima
arima_model <- auto.arima(train_data, seasonal = FALSE)
# Fitting series
Box.test(arima_model$fitted[,1], lag = 24, type = "Ljung-Box")
adf.test(arima_model$fitted[,1], alternative = "stationary")
# Fitting residuals
Box.test(arima_model$residuals, lag = 24, type = "Ljung-Box")
ks.test(scale(arima_model$residuals), "pnorm")

# SArima
sarima_model <- auto.arima(train_data, seasonal = TRUE)
# Fitting series
Box.test(sarima_model$fitted[,1], lag = 24, type = "Ljung-Box")
adf.test(sarima_model$fitted[,1], alternative = "stationary")
# Fitting residuals
Box.test(sarima_model$residuals, lag = 24, type = "Ljung-Box")
ks.test(scale(sarima_model$residuals), "pnorm")

# ETS
ets_model <- ets(train_data)
# Fitting series
Box.test(ets_model$fitted[,1], lag = 24, type = "Ljung-Box")
adf.test(ets_model$fitted[,1], alternative = "stationary")
# Fitting residuals
Box.test(ets_model$residuals, lag = 24, type = "Ljung-Box")
ks.test(scale(ets_model$residuals), "pnorm")

# Holt-Winters
holtwinters_model <- HoltWinters(train_data)
holtwinters_forecast <- forecast(holtwinters_model, h = length(test_data), level = 95)
# Fitting series
Box.test(holtwinters_model$fitted[,1], lag = 24, type = "Ljung-Box")
adf.test(holtwinters_model$fitted[,1], alternative = "stationary")
# Fitting residuals
Box.test(holtwinters_forecast$residuals[13:72], lag = 24, type = "Ljung-Box")
ks.test(scale(holtwinters_forecast$residuals[13:72]), "pnorm")

# TBATS
tbats_model <- tbats(train_data)
# Fitting series
Box.test(tbats_model$fitted.values, lag = 24, type = "Ljung-Box")
adf.test(tbats_model$fitted.values, alternative = "stationary")
# Fitting residuals
Box.test(tbats_model$errors, lag = 24, type = "Ljung-Box")
ks.test(scale(tbats_model$errors), "pnorm")

# STSM
bsm_model <- StructTS(train_data, type = "BSM")
# Fitting series
Box.test(bsm_model$fitted[,1], lag = 24, type = "Ljung-Box")
adf.test(bsm_model$fitted[,1], alternative = "stationary")
# Fitting residuals
bsm_model_residuals <- train_data - bsm_model$fitted[,1]
Box.test(bsm_model_residuals, lag = 24, type = "Ljung-Box")
ks.test(scale(bsm_model_residuals), "pnorm")

# TSLM
tslm_model <- tslm(train_data ~ trend + season)
# Fitting series
Box.test(tslm_model$fitted.values, lag = 24, type = "Ljung-Box")
adf.test(tslm_model$fitted.values, alternative = "stationary")
# Fitting residuals
Box.test(tslm_model$residuals, lag = 24, type = "Ljung-Box")
ks.test(scale(tslm_model$residuals), "pnorm")

# prophet
date_prophet <- seq(as.Date("2013-01-01"), as.Date("2018-12-01"), by = "month")
train_prophet <- data.frame(ds = date_prophet, y = as.numeric(train_data))
prophet_model <- prophet(train_prophet, growth = "linear", seasonality.mode =  "additive", changepoint.prior.scale = 0.03)
future <- make_future_dataframe(prophet_model, period = 12, freq = "month")
prophet_predict <- predict(prophet_model, future)
prophet_model$fitted <- ts(prophet_predict$yhat[1:72], frequency = 12, start = c(2013,1))
# Fitting series
Box.test(prophet_model$fitted, lag = 24, type = "Ljung-Box")
adf.test(prophet_model$fitted, alternative = "stationary")
# Fitting residuals
prophet_model_residuals <- train_data - prophet_model$fitted
Box.test(prophet_model_residuals, lag = 24, type = "Ljung-Box")
ks.test(scale(prophet_model_residuals), "pnorm")

# NNAR
nnar_model <- nnetar(train_data, repeats = 50, decay = 0.15)
# Fitting series
Box.test(nnar_model$fitted[13:72], lag = 24, type = "Ljung-Box")
adf.test(nnar_model$fitted[13:72], alternative = "stationary")
# Fitting residuals
Box.test(nnar_model$residuals[13:72], lag = 24, type = "Ljung-Box")
ks.test(scale(nnar_model$residuals[13:72]), "pnorm")
```

#2.Table 3
```{r}
# raw data
ts_data <- openxlsx::read.xlsx("./raw data for table 2:3.xlsx", sheet = 1, rows = 38:160, cols = 3)
ts_data <- ts(ts_data, frequency = 12, start = c(2013,1))
train_data <- window(ts_data, end = c(2019,12))
test_data <- window(ts_data, start = c(2020,1), end = c(2023,2))

# Sarima
sarima_model <- auto.arima(train_data, seasonal = TRUE)
sarima_forecast <- forecast(sarima_model, h = length(test_data), level = 95)

# Holt-Winters
holtwinters_model <- HoltWinters(train_data)
holtwinters_forecast <- forecast(holtwinters_model, h = length(test_data), level = 95)

# NNAR
nnar_model <- nnetar(train_data, repeats = 50, decay = 0.15)
nnar_forecast <- forecast(nnar_model, h=length(test_data), PI = TRUE, npaths = 150, level = 95)

# 2020.1-2020.5:[1:5]
# 2020.6-2021.2:[6:14]
# 2021.3-2022.11:[15:35]
# 2022.12-2023.1:[36:37]
# 2023.2:[38]
model_period_error <- 
data.frame(
  sarima = c( 
    round(Mape(sarima_forecast$mean[1:5], test_data[1:5]),4),
    round(Mape(sarima_forecast$mean[6:14], test_data[6:14]),4),
    round(Mape(sarima_forecast$mean[15:35], test_data[15:35]),4), 
    round(Mape(sarima_forecast$mean[36:37], test_data[36:37]),4),
    round(Mape(sarima_forecast$mean[38], test_data[38]),4),
    round(Rmse(sarima_forecast$mean[1:5], test_data[1:5]),4),
    round(Rmse(sarima_forecast$mean[6:14], test_data[6:14]),4),
    round(Rmse(sarima_forecast$mean[15:35], test_data[15:35]),4), 
    round(Rmse(sarima_forecast$mean[36:37], test_data[36:37]),4),
    round(Rmse(sarima_forecast$mean[38], test_data[38]),4)
            ),
  holtwinters = c( 
    round(Mape(holtwinters_forecast$mean[1:5], test_data[1:5]),4),
    round(Mape(holtwinters_forecast$mean[6:14], test_data[6:14]),4),
    round(Mape(holtwinters_forecast$mean[15:35], test_data[15:35]),4), 
    round(Mape(holtwinters_forecast$mean[36:37], test_data[36:37]),4),
    round(Mape(holtwinters_forecast$mean[38], test_data[38]),4),
    round(Rmse(holtwinters_forecast$mean[1:5], test_data[1:5]),4),
    round(Rmse(holtwinters_forecast$mean[6:14], test_data[6:14]),4),
    round(Rmse(holtwinters_forecast$mean[15:35], test_data[15:35]),4), 
    round(Rmse(holtwinters_forecast$mean[36:37], test_data[36:37]),4),
    round(Rmse(holtwinters_forecast$mean[38], test_data[38]),4)
               ),
  nnar = c( 
    round(Mape(nnar_forecast$mean[1:5], test_data[1:5]),4),
    round(Mape(nnar_forecast$mean[6:14], test_data[6:14]),4),
    round(Mape(nnar_forecast$mean[15:35], test_data[15:35]),4), 
    round(Mape(nnar_forecast$mean[36:37], test_data[36:37]),4),
    round(Mape(nnar_forecast$mean[38], test_data[38]),4),
    round(Rmse(nnar_forecast$mean[1:5], test_data[1:5]),4),
    round(Rmse(nnar_forecast$mean[6:14], test_data[6:14]),4),
    round(Rmse(nnar_forecast$mean[15:35], test_data[15:35]),4), 
    round(Rmse(nnar_forecast$mean[36:37], test_data[36:37]),4),
    round(Rmse(nnar_forecast$mean[38], test_data[38]),4)
            ),
  row.names = c("mape:2020Jan-2020May","mape:2020Jun-2021Feb",
                "mape:2021Mar-2022Nov","mape:2022Dec-2023Jan","mape:2023Feb",
                "rmse:2020Jan-2020May","rmse:2020Jun-2021Feb",
                "rmse:2021Mar-2022Nov","rmse:2022Dec-2023Jan","rmse:2023Feb")
  )
print(model_period_error)
```

